---
title: "CUPED Simulation: Hulu A/B Test"
author: "ECMA 31370 White Paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

## 1. Overview

We simulate an A/B test on Hulu testing a new recommendation algorithm. The primary metric is weekly streaming hours per user. The goal is to compare four estimators for the treatment effect and demonstrate how CUPED reduces variance.

## 2. Data Generating Process

Each user $i$ has a stable baseline streaming rate $\mu_i$ drawn from a lognormal distribution. Weekly streaming hours are:

$$\text{hours}_{it} = \max\left(0,\; \mu_i + \tau \cdot W_i + \varepsilon_{it}\right)$$

where $W_i \in \{0,1\}$ is the treatment indicator, $\tau = 0.10$ hrs/week is the true treatment effect, and $\varepsilon_{it} \sim N(0, 5^2)$ is week-to-week noise. The floor at zero ensures non-negative hours. Light users naturally produce zero-hours weeks when noise pushes them below zero, and because the same users tend to be light across all weeks, these zeros are correlated across pre and post periods.

```{r parameters}
set.seed(42)

N <- 50000               # total users
new_user_frac <- 0.20    # 20% new users (no history)
log_mean <- log(7)       # baseline median ~7 hrs/week
log_sd <- 0.7            # right-skewed spread
noise_sd <- 5.0          # week-to-week noise (tuned for rho ~ 0.70)
tau <- 0.10              # treatment effect: +0.10 hrs/week
n_pre_weeks <- 4         # 4 weeks of pre-experiment data
```

| Parameter | Value | Rationale |
|-----------|-------|-----------|
| N | 50,000 | Mid-tier platform experiment scale |
| New user fraction | 20% | Growing platform |
| Baseline $\mu_i$ | LogNormal(log 7, 0.7) | Median 7 hrs/week, right-skewed to 30+ |
| Noise $\sigma_\varepsilon$ | 5.0 | Tuned so $\rho \approx 0.70$ for 1-week window |
| Treatment $\tau$ | 0.10 hrs/week | Small enough that naive method struggles |

## 3. Generate Users

```{r generate-users}
n_new <- round(N * new_user_frac)
n_existing <- N - n_new

users <- tibble(
  user_id = 1:N,
  is_new = c(rep(TRUE, n_new), rep(FALSE, n_existing)),
  baseline_mu = rlnorm(N, meanlog = log_mean, sdlog = log_sd),
  treatment = sample(c(0L, 1L), N, replace = TRUE)
)

cat(sprintf("Treatment: %d | Control: %d\n",
            sum(users$treatment == 1), sum(users$treatment == 0)))
cat(sprintf("New: %d | Existing: %d\n", n_new, n_existing))
```

## 4. Generate Weekly Streaming Hours

```{r generate-weeks}
generate_week <- function(baseline_mu, treatment_indicator = NULL, tau_val = 0) {
  n <- length(baseline_mu)
  noise <- rnorm(n, mean = 0, sd = noise_sd)
  hours <- baseline_mu + noise
  if (!is.null(treatment_indicator)) {
    hours <- hours + tau_val * treatment_indicator
  }
  return(pmax(hours, 0))
}

pre_weeks <- map(1:n_pre_weeks, ~ generate_week(users$baseline_mu))
names(pre_weeks) <- paste0("pre_week_", 1:n_pre_weeks)

post_week <- generate_week(users$baseline_mu, users$treatment, tau)
```

## 5. Assemble Dataset

Three pre-experiment covariates at different window lengths. New users get `NA`.

```{r assemble}
sim_data <- users %>%
  bind_cols(as_tibble(pre_weeks)) %>%
  mutate(
    post_week = post_week,
    pre_1wk = pre_week_4,
    pre_2wk = (pre_week_3 + pre_week_4) / 2,
    pre_4wk = (pre_week_1 + pre_week_2 + pre_week_3 + pre_week_4) / 4
  ) %>%
  mutate(
    pre_1wk = if_else(is_new, NA_real_, pre_1wk),
    pre_2wk = if_else(is_new, NA_real_, pre_2wk),
    pre_4wk = if_else(is_new, NA_real_, pre_4wk)
  )

head(sim_data %>% select(user_id, is_new, treatment, pre_1wk, pre_4wk, post_week), 10)
```

## 6. Explore the Data

### 6.1 Distribution of Streaming Hours

```{r dist-plot, fig.width=8, fig.height=4}
ggplot(sim_data, aes(x = post_week)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white", alpha = 0.8) +
  labs(title = "Distribution of Weekly Streaming Hours (Experiment Period)",
       x = "Streaming Hours / Week", y = "Count") +
  theme_minimal()
```

```{r dist-summary}
cat(sprintf("Mean:   %.2f hrs/week\n", mean(sim_data$post_week)))
cat(sprintf("Median: %.2f hrs/week\n", median(sim_data$post_week)))
cat(sprintf("SD:     %.2f\n", sd(sim_data$post_week)))
cat(sprintf("Zero-hours users: %.1f%%\n", mean(sim_data$post_week == 0) * 100))
```

### 6.2 Pre-Post Correlation

```{r pre-post-corr}
existing <- sim_data %>% filter(!is_new)

rho_1 <- cor(existing$pre_1wk, existing$post_week)
rho_2 <- cor(existing$pre_2wk, existing$post_week)
rho_4 <- cor(existing$pre_4wk, existing$post_week)

tibble(
  Window = c("1 week", "2 weeks", "4 weeks"),
  rho = round(c(rho_1, rho_2, rho_4), 3),
  `Theoretical VR` = paste0(round(c(rho_1, rho_2, rho_4)^2 * 100, 1), "%")
) %>% knitr::kable()
```

### 6.3 Pre vs Post Scatter Plot

```{r scatter, fig.width=8, fig.height=4}
existing %>%
  sample_n(5000) %>%
  ggplot(aes(x = pre_4wk, y = post_week)) +
  geom_point(alpha = 0.15, size = 0.8, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 0.8) +
  labs(title = sprintf("Pre vs Post Streaming Hours (4-week window, rho = %.3f)", rho_4),
       x = "Pre-Experiment (4-week avg, hrs/week)",
       y = "Experiment Period (hrs/week)") +
  theme_minimal()
```

## 7. CUPED Implementation

$$Y_i^{cv} = Y_i - \hat{\theta}(X_i - \bar{X})$$

where $\hat{\theta} = \text{Cov}(Y, X) / \text{Var}(X)$. For users with missing $X_i$ (new users), we apply the Booking.com COALESCE approach: leave their metric unadjusted.

```{r cuped-function}
cuped_adjust <- function(y, x) {
  has_x <- !is.na(x)
  theta <- cov(y[has_x], x[has_x]) / var(x[has_x])
  x_mean <- mean(x[has_x])
  y_adj <- y
  y_adj[has_x] <- y[has_x] - theta * (x[has_x] - x_mean)
  return(list(y_adj = y_adj, theta = theta))
}
```

## 8. Four-Method Comparison

We compare: Naive DM, First Differences ($\theta = 1$), CUPED (optimal $\hat{\theta}$), and OLS Regression Adjustment.

```{r run-test-all}
run_test_all <- function(data, pre_col, label) {
  y <- data$post_week
  x <- data[[pre_col]]
  w <- data$treatment
  has_x <- !is.na(x)

  y_t <- y[w == 1]; y_c <- y[w == 0]
  n_t <- length(y_t); n_c <- length(y_c)

  # Naive DM
  naive_est <- mean(y_t) - mean(y_c)
  naive_se <- sqrt(var(y_t) / n_t + var(y_c) / n_c)
  naive_ci <- naive_est + c(-1, 1) * 1.96 * naive_se

  # First Differences (theta = 1)
  fd <- y; fd[has_x] <- y[has_x] - x[has_x]
  fd_t <- fd[w == 1]; fd_c <- fd[w == 0]
  fd_est <- mean(fd_t) - mean(fd_c)
  fd_se <- sqrt(var(fd_t) / n_t + var(fd_c) / n_c)
  fd_ci <- fd_est + c(-1, 1) * 1.96 * fd_se

  # CUPED (optimal theta)
  res <- cuped_adjust(y, x)
  ya <- res$y_adj
  ya_t <- ya[w == 1]; ya_c <- ya[w == 0]
  cuped_est <- mean(ya_t) - mean(ya_c)
  cuped_se <- sqrt(var(ya_t) / n_t + var(ya_c) / n_c)
  cuped_ci <- cuped_est + c(-1, 1) * 1.96 * cuped_se

  # OLS Regression Adjustment
  df_ols <- tibble(y = y, w = w, x = x) %>% filter(!is.na(x))
  mod <- lm(y ~ w + x, data = df_ols)
  ols_est <- coef(mod)["w"]
  ols_se <- summary(mod)$coefficients["w", "Std. Error"]
  ols_ci <- ols_est + c(-1, 1) * 1.96 * ols_se

  sig <- function(ci) ifelse((ci[1] > 0) | (ci[2] < 0), "Yes", "No")

  tibble(
    Window = label,
    Method = c("Naive DM", "First Diff.", "CUPED", "OLS Reg. Adj."),
    Estimate = round(c(naive_est, fd_est, cuped_est, ols_est), 4),
    SE = round(c(naive_se, fd_se, cuped_se, ols_se), 4),
    `95% CI` = c(sprintf("[%.4f, %.4f]", naive_ci[1], naive_ci[2]),
                 sprintf("[%.4f, %.4f]", fd_ci[1], fd_ci[2]),
                 sprintf("[%.4f, %.4f]", cuped_ci[1], cuped_ci[2]),
                 sprintf("[%.4f, %.4f]", ols_ci[1], ols_ci[2])),
    Sig. = c(sig(naive_ci), sig(fd_ci), sig(cuped_ci), sig(ols_ci)),
    `Var. Red.` = sprintf("%.1f%%", c(0,
                   (1 - fd_se^2 / naive_se^2) * 100,
                   (1 - cuped_se^2 / naive_se^2) * 100,
                   (1 - ols_se^2 / naive_se^2) * 100))
  )
}
```

### 8.1 Existing Users

```{r results-existing}
all_results <- bind_rows(
  run_test_all(existing, "pre_1wk", "1-week"),
  run_test_all(existing, "pre_2wk", "2-week"),
  run_test_all(existing, "pre_4wk", "4-week")
)

all_results %>% knitr::kable()
```

### 8.2 All Users (COALESCE approach)

```{r results-all}
bind_rows(
  run_test_all(sim_data, "pre_1wk", "1-week"),
  run_test_all(sim_data, "pre_2wk", "2-week"),
  run_test_all(sim_data, "pre_4wk", "4-week")
) %>% knitr::kable()
```

### 8.3 New Users Only

```{r results-new}
new_users <- sim_data %>% filter(is_new)
y_t <- new_users$post_week[new_users$treatment == 1]
y_c <- new_users$post_week[new_users$treatment == 0]
est <- mean(y_t) - mean(y_c)
se <- sqrt(var(y_t) / length(y_t) + var(y_c) / length(y_c))
ci <- est + c(-1, 1) * 1.96 * se

tibble(
  Estimate = est,
  SE = se,
  `95% CI` = sprintf("[%.4f, %.4f]", ci[1], ci[2]),
  Significant = ifelse((ci[1] > 0) | (ci[2] < 0), "Yes", "No")
) %>% knitr::kable()
```

No CUPED adjustment is possible for new users. The SE is roughly 2x that of existing users.

## 9. Confidence Interval Comparison

```{r ci-plot, fig.width=8, fig.height=5}
plot_data <- all_results %>%
  mutate(
    CI_lo = as.numeric(str_extract(`95% CI`, "(?<=\\[)[\\d.-]+")),
    CI_hi = as.numeric(str_extract(`95% CI`, "(?<=, )[\\d.-]+(?=\\])"))
  ) %>%
  rename(Est = Estimate)

plot_data$Window <- factor(plot_data$Window, levels = c("1-week", "2-week", "4-week"))
plot_data$Method <- factor(plot_data$Method, levels = c("Naive DM", "First Diff.", "CUPED", "OLS Reg. Adj."))

ggplot(plot_data, aes(x = Window, y = Est, color = Method)) +
  geom_pointrange(aes(ymin = CI_lo, ymax = CI_hi),
                  position = position_dodge(width = 0.5), size = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_hline(yintercept = tau, linetype = "dotted", color = "red", alpha = 0.6) +
  annotate("text", x = 2.8, y = tau + 0.02, label = sprintf("True tau = %.2f", tau),
           color = "red", size = 3) +
  scale_color_manual(values = c("Naive DM" = "gray50", "First Diff." = "orange3",
                                "CUPED" = "steelblue", "OLS Reg. Adj." = "darkgreen")) +
  labs(title = "Four-Method Comparison: 95% Confidence Intervals (Existing Users)",
       y = "Estimated Treatment Effect (hrs/week)",
       x = "Pre-Experiment Window Length") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
