---
title: "ECMA 31370, CUPED White Paper"
author: "YOUR NAME, University of Chicago"
geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
date: "XX-XX-2025"
output: pdf_document
header-includes:
  - \usepackage{caption}
  - \usepackage{float}
  - \PassOptionsToPackage{dvipsnames}{xcolor}
  - \usepackage{amsmath}
  - \usepackage{enumitem}
  - \usepackage{soul}
  - \usepackage[skins,breakable]{tcolorbox}
  - \usepackage{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H", out.extra = "")
```

# Executive Summary

---

# White Paper

## [1] State the Problem and Ask

Hulu is a subscription-based video streaming service that runs A/B tests to evaluate product changes such as recommendation algorithms, UI redesigns, and content curation strategies. All experiments are currently analyzed using difference-in-means (DM). The problem is that most product changes produce small lifts (1 to 3% relative to baseline), and user-level streaming behavior is highly variable, with weekly hours ranging from zero to 30+. The noise in the data regularly overwhelms the treatment signal, so experiments that produce real improvements often end up with confidence intervals that include zero. Kohavi and Thomke (2017) document that this is a common pattern across the industry: variance, not sample size or idea quality, is the binding constraint in most online experiments.

We ask whether adopting CUPED (Controlled-experiment Using Pre-Experiment Data) can solve this problem. Our proof-of-concept simulation (Section 7) shows that it can: with Hulu's data characteristics, the naive DM estimator fails to detect a true 0.10 hrs/week treatment effect, while CUPED with a 2-week or longer pre-experiment window detects it.

## [2] Make a Recommendation

We recommend that Hulu adopt CUPED as the default analysis method for all A/B tests. The covariate should be pre-experiment weekly streaming hours (which Hulu already logs), with a default window of 4 weeks. For new users without history (about 20% of the experiment population), we follow Booking.com's COALESCE approach and leave their metrics unadjusted (Jackson, 2018). CUPED should be built into the analysis pipeline as a default rather than an opt-in tool, as Meta has done across hundreds of thousands of concurrent experiments (Meakin & Sangwan, 2024).

Our simulation (Section 7) and industry benchmarks suggest 40 to 60% variance reduction for existing users. Booking.com reported reaching significance roughly twice as fast after adopting CUPED (Jackson, 2018). For Hulu, this translates to either detecting smaller effects at the same sample size, or reaching conclusions in half the time.

## [3] Describe the CUPED Methodology

CUPED (Controlled-experiment Using Pre-Experiment Data) was introduced by Deng et al. (2013) at Microsoft. The core idea borrows from the control variates technique in Monte Carlo simulation: if we observe a variable $X$ that is correlated with the outcome $Y$ but unaffected by the treatment, we can subtract out the predictable component of $Y$ to obtain a lower-variance estimator without introducing bias.

Concretely, let $Y_i$ be the experiment-period outcome for user $i$, $X_i$ be a pre-experiment covariate (typically the same metric measured before the experiment started), and $W_i \in \{0,1\}$ be the treatment indicator. The CUPED-adjusted outcome is:

$$Y_i^{cv} = Y_i - \hat{\theta}(X_i - \bar{X})$$

where $\hat{\theta} = \text{Cov}(Y, X) / \text{Var}(X)$ is estimated from the data. The subtracted term $\hat{\theta}(X_i - \bar{X})$ removes the portion of $Y_i$ that is predictable from the pre-experiment covariate. Because treatment is randomly assigned and therefore independent of $X$, this adjustment does not change the expected value of the treatment effect estimator. The treatment effect is then estimated as the difference in means of $Y^{cv}$ between treatment and control groups, just as in the standard approach.

The variance of the adjusted estimator is reduced by a factor of $1 - \rho^2$, where $\rho = \text{Cor}(Y, X)$. A correlation of $\rho = 0.7$ yields about 50% variance reduction; $\rho = 0.8$ yields about 64%. This means the pre-post correlation is the single most important quantity determining whether CUPED is worthwhile for a given metric. Metrics with stable user-level behavior across time (such as streaming hours, purchase frequency, or session count) tend to have high $\rho$ and benefit most. Metrics that are inherently volatile or one-time in nature benefit less.

CUPED is applied after the experiment is complete, using only data that was observed before randomization. This makes it compatible with any randomization scheme and easy to add to an existing analysis pipeline without changing how experiments are designed or run.

## [4] Compare CUPED to Other Methodologies

The standard approach to analyzing an A/B test is difference-in-means (DM): compute the average outcome in the treatment group, subtract the average in the control group, and construct a confidence interval using the pooled standard error. DM is unbiased under random assignment, but it ignores all pre-experiment information, so its variance reflects the full heterogeneity in user-level outcomes. When the metric is noisy (as streaming hours are), DM often fails to detect small but real treatment effects.

First differences is the simplest way to incorporate pre-experiment data. For each user, compute $\Delta_i = Y_{i,post} - X_{i,pre}$, then take the difference in means of $\Delta$ between treatment and control. This is algebraically equivalent to CUPED with $\theta$ forced to 1. When the pre and post metrics are measured in the same units and have similar variance, $\theta$ is close to 1 and first differences performs nearly as well as CUPED. But when the relationship between pre and post is not one-to-one (for example, when the pre-experiment covariate is an average over multiple weeks, or when the outcome distribution has a floor at zero), the optimal $\theta$ deviates from 1 and first differences leaves variance on the table. In our simulation, $\hat{\theta}$ ranges from 0.69 (1-week window) to 0.90 (4-week window), and the gap in variance reduction between first differences and CUPED is largest at short windows.

OLS regression adjustment fits $Y_i = \alpha + \beta W_i + \gamma X_i + \varepsilon_i$ and reads the treatment effect from $\hat{\beta}$. With a single covariate, this produces exactly the same estimate and standard error as CUPED, because the OLS coefficient $\hat{\gamma}$ equals CUPED's $\hat{\theta}$. The equivalence breaks when multiple covariates are included: OLS generalizes naturally to a vector of regressors, while single-covariate CUPED does not. Deng et al. (2022) at Airbnb formalized this multi-covariate extension as CUPAC, showing that variance reduction equals $R^2$ from regressing the outcome on the full covariate vector.

Difference-in-differences (DiD) is designed for settings where treatment and control groups may have different pre-existing trends, such as when randomization is at the market or region level rather than the user level. DiD computes the change in the treatment group's mean minus the change in the control group's mean, under the parallel trends assumption. In a standard user-level A/B test with proper randomization, the parallel trends assumption is satisfied by construction and DiD reduces to first differences. DiD is therefore most useful in quasi-experimental settings where randomization is imperfect or unavailable, rather than in the fully randomized experiments that CUPED is designed for.

The choice among these methods depends on the experimental setting. For a properly randomized user-level A/B test like the ones Hulu runs, CUPED (or equivalently, OLS with a single pre-experiment covariate) is the most practical option: it extracts the maximum variance reduction from the available pre-experiment data, requires no changes to the randomization procedure, and is straightforward to implement. First differences is a reasonable default when the analyst does not want to estimate $\theta$, but it will underperform CUPED when $\theta$ differs from 1. DM should be reserved for cases where no pre-experiment data is available.

## [5] Report on CUPED Adoption Elsewhere

Netflix was among the first to evaluate CUPED in production. Xie and Aurisset (2016) compared it against stratified sampling and post-stratification using previous-week streaming hours as the covariate. For existing users with viewing history, all three methods achieved roughly 40% variance reduction at $\rho = 0.7$ to $0.8$. Stratified sampling, however, degraded from about 45% to 15% in Netflix's distributed multi-machine environment because balanced assignment could not be maintained across servers; CUPED, applied after data collection, was unaffected. Netflix concluded that post-assignment methods are preferable and that the choice of covariate matters more than the choice of method. They also found that new users without viewing history offered little to adjust on: $\rho$ dropped to $0.2$--$0.4$ and variance reduction was negligible.

Booking.com (Jackson, 2018) tackled the missing-data problem that Netflix flagged. Booking.com processes over 1.5 million room-night reservations per day, but with a conversion rate of about 2%, detecting even a 1% relative improvement requires over 15 million users under standard methods, so variance reduction is especially valuable at their scale. Rather than excluding first-time visitors who lack pre-experiment data, they implemented a one-line SQL COALESCE pattern that leaves those users' metrics unadjusted. Mathematically, this is equivalent to imputing missing covariates with the sample mean, which zeroes out the adjustment term. In a live experiment on a new calendar interface, the CUPED-adjusted metric reached significance around day 5--6, while the unadjusted metric did not reach significance until day 11--12, roughly halving the time needed to call the experiment. The COALESCE approach has since become standard practice at other companies as well.

Faire (Ritchie & Wong, 2021), a B2B wholesale marketplace with sample sizes in the tens of thousands, ran the most systematic study of pre-experiment window length. They tested windows of 1, 2, 4, 8, and 16 weeks: variance reduction rose from about 22% at 1 week to 45% at 8 weeks and plateaued at 50% by 16 weeks. Faire settled on 8 weeks as their default, noting that 90% of consecutive orders on their platform occur within 28 days. They also combined CUPED with outlier capping at 5 standard deviations and found the two techniques complementary, since outlier capping limits the influence of extreme values while CUPED removes baseline user-level variation.

DoorDash (Zhang & Tang, 2020) illustrates a different dimension of CUPED's value. Before adopting variance reduction, the platform could run only about 2 concurrent experiments on its consumer app because high variance in order-value metrics required large sample sizes that consumed most of the available traffic. After implementing CUPED alongside experiment parallelization, they achieved a 4X increase in experiment capacity. DoorDash also found that the type of covariate matters: user-level behavioral data (such as order history) reduced the required sample size by 10--20%, while categorical attributes like platform (iOS vs Android) and market (city) contributed only 2--4%. The lesson is that investing in user-level pre-experiment features pays off far more than relying on broad demographic segments.

CUPED has since scaled well beyond these early adopters. Meta applies it across hundreds of thousands of concurrent experiments under the internal name "Mean2.0," with over 30% average confidence interval reduction (Meakin & Sangwan, 2024). Commercial platforms including Statsig and Eppo now ship CUPED as an enabled-by-default feature (Sexauer, 2024).

These experiences inform several design choices for Hulu. Netflix used weekly streaming hours as its core metric and observed $\rho \approx 0.7$ with ~40% variance reduction, which provides a direct benchmark for what Hulu can expect. The new-user limitation Netflix documented applies as well, given that roughly 20% of Hulu's experiment participants are new subscribers without viewing history. Booking.com's COALESCE approach offers a ready-made solution for including these users without discarding them. Faire's window-length analysis suggests that a 4-week default is a natural starting point for a platform where viewing habits follow a weekly rhythm, and that extending beyond 4 weeks yields diminishing returns. DoorDash's finding on covariate quality reinforces the case for using streaming hours (a user-level behavioral metric) rather than demographic segments as the primary covariate. Our simulation in Section 7 tests these choices under Hulu-like parameters and produces results consistent with the patterns reported above.

## [6] Describe Implementation

## [7] Report on a POC

We simulate an A/B test on Hulu to illustrate how CUPED reduces estimator variance. The platform is testing a new recommendation algorithm, and the primary metric is weekly streaming hours per user. Full simulation code is provided in Appendix A.

### Data Generating Process

Each user $i$ has a stable baseline streaming rate $\mu_i \sim \text{LogNormal}(\log 7, 0.7)$, producing a right-skewed distribution with median around 7 hours per week. Weekly streaming hours are generated as:

$$\text{hours}_{it} = \max\left(0,\; \mu_i + \tau \cdot W_i + \varepsilon_{it}\right)$$

where $W_i \in \{0,1\}$ is the treatment indicator, $\tau = 0.10$ hours/week is the true treatment effect, and $\varepsilon_{it} \sim N(0, 4^2)$ is independent week-to-week noise. The floor at zero ensures non-negative hours. Light users with small $\mu_i$ naturally produce zero-hours weeks when noise pushes them below zero, and because the same users tend to be light across all weeks, these zeros are correlated across pre and post periods, preserving realistic pre-post correlation without any additional mechanism.

We generate 50,000 users (80% existing, 20% new), 4 weeks of pre-experiment data, and 1 week of experiment data. New users have no pre-experiment observations.

```{r poc-setup, include=FALSE}
library(tidyverse)
```

```{r poc-generate, echo=FALSE}
set.seed(132)

N <- 50000
new_user_frac <- 0.20
log_mean <- log(7)
log_sd <- 0.7
noise_sd <- 4.0
tau <- 0.10
n_pre_weeks <- 4

n_new <- round(N * new_user_frac)
n_existing <- N - n_new

users <- tibble(
  user_id = 1:N,
  is_new = c(rep(TRUE, n_new), rep(FALSE, n_existing)),
  baseline_mu = rlnorm(N, meanlog = log_mean, sdlog = log_sd),
  treatment = sample(c(0L, 1L), N, replace = TRUE)
)

generate_week <- function(baseline_mu, treatment_indicator = NULL, tau_val = 0) {
  n <- length(baseline_mu)
  noise <- rnorm(n, mean = 0, sd = noise_sd)
  hours <- baseline_mu + noise
  if (!is.null(treatment_indicator)) {
    hours <- hours + tau_val * treatment_indicator
  }
  return(pmax(hours, 0))
}

pre_weeks <- map(1:n_pre_weeks, ~ generate_week(users$baseline_mu))
names(pre_weeks) <- paste0("pre_week_", 1:n_pre_weeks)
post_week <- generate_week(users$baseline_mu, users$treatment, tau)

sim_data <- users %>%
  bind_cols(as_tibble(pre_weeks)) %>%
  mutate(
    post_week = post_week,
    pre_1wk = pre_week_4,
    pre_2wk = (pre_week_3 + pre_week_4) / 2,
    pre_4wk = (pre_week_1 + pre_week_2 + pre_week_3 + pre_week_4) / 4
  ) %>%
  mutate(
    pre_1wk = if_else(is_new, NA_real_, pre_1wk),
    pre_2wk = if_else(is_new, NA_real_, pre_2wk),
    pre_4wk = if_else(is_new, NA_real_, pre_4wk)
  )

existing <- sim_data %>% filter(!is_new)

cuped_adjust <- function(y, x) {
  has_x <- !is.na(x)
  theta <- cov(y[has_x], x[has_x]) / var(x[has_x])
  x_mean <- mean(x[has_x])
  y_adj <- y
  y_adj[has_x] <- y[has_x] - theta * (x[has_x] - x_mean)
  return(list(y_adj = y_adj, theta = theta))
}

run_test_all <- function(data, pre_col, label) {
  y <- data$post_week
  x <- data[[pre_col]]
  w <- data$treatment
  has_x <- !is.na(x)

  y_t <- y[w == 1]; y_c <- y[w == 0]
  n_t <- length(y_t); n_c <- length(y_c)

  # Naive DM
  naive_est <- mean(y_t) - mean(y_c)
  naive_se <- sqrt(var(y_t) / n_t + var(y_c) / n_c)
  naive_ci <- naive_est + c(-1, 1) * 1.96 * naive_se

  # First Differences (theta = 1)
  fd <- y; fd[has_x] <- y[has_x] - x[has_x]
  fd_t <- fd[w == 1]; fd_c <- fd[w == 0]
  fd_est <- mean(fd_t) - mean(fd_c)
  fd_se <- sqrt(var(fd_t) / n_t + var(fd_c) / n_c)
  fd_ci <- fd_est + c(-1, 1) * 1.96 * fd_se

  # CUPED (optimal theta)
  res <- cuped_adjust(y, x)
  ya <- res$y_adj
  ya_t <- ya[w == 1]; ya_c <- ya[w == 0]
  cuped_est <- mean(ya_t) - mean(ya_c)
  cuped_se <- sqrt(var(ya_t) / n_t + var(ya_c) / n_c)
  cuped_ci <- cuped_est + c(-1, 1) * 1.96 * cuped_se

  # OLS Regression Adjustment
  df_ols <- tibble(y = y, w = w, x = x) %>% filter(!is.na(x))
  mod <- lm(y ~ w + x, data = df_ols)
  ols_est <- coef(mod)["w"]
  ols_se <- summary(mod)$coefficients["w", "Std. Error"]
  ols_ci <- ols_est + c(-1, 1) * 1.96 * ols_se

  sig <- function(ci) ifelse((ci[1] > 0) | (ci[2] < 0), "Yes", "No")

  tibble(
    Window = label,
    Method = c("Naive DM", "First Diff.", "CUPED", "OLS Reg. Adj."),
    Estimate = round(c(naive_est, fd_est, cuped_est, ols_est), 4),
    SE = round(c(naive_se, fd_se, cuped_se, ols_se), 4),
    `95% CI` = c(sprintf("[%.4f, %.4f]", naive_ci[1], naive_ci[2]),
                 sprintf("[%.4f, %.4f]", fd_ci[1], fd_ci[2]),
                 sprintf("[%.4f, %.4f]", cuped_ci[1], cuped_ci[2]),
                 sprintf("[%.4f, %.4f]", ols_ci[1], ols_ci[2])),
    Sig. = c(sig(naive_ci), sig(fd_ci), sig(cuped_ci), sig(ols_ci)),
    `Var. Red.` = sprintf("%.1f%%", c(0,
                   (1 - fd_se^2 / naive_se^2) * 100,
                   (1 - cuped_se^2 / naive_se^2) * 100,
                   (1 - ols_se^2 / naive_se^2) * 100))
  )
}
```

### Pre-Post Correlation and Method Comparison

```{r poc-compute, echo=FALSE}
rho_1 <- cor(existing$pre_1wk, existing$post_week)
rho_2 <- cor(existing$pre_2wk, existing$post_week)
rho_4 <- cor(existing$pre_4wk, existing$post_week)

all_results <- bind_rows(
  run_test_all(existing, "pre_1wk", "1-week"),
  run_test_all(existing, "pre_2wk", "2-week"),
  run_test_all(existing, "pre_4wk", "4-week")
)

# all users, 4-week CUPED
res_all <- run_test_all(sim_data, "pre_4wk", "4-week")
cuped_all_vr <- res_all %>% filter(Method == "CUPED") %>% pull(`Var. Red.`)

# new users naive
new_users <- sim_data %>% filter(is_new)
y_t <- new_users$post_week[new_users$treatment == 1]
y_c <- new_users$post_week[new_users$treatment == 0]
new_se <- round(sqrt(var(y_t) / length(y_t) + var(y_c) / length(y_c)), 4)
ex_naive_se <- round(res_all %>% filter(Method == "Naive DM") %>% pull(SE), 4)

cuped_vr_1 <- all_results %>% filter(Method == "CUPED", Window == "1-week") %>% pull(`Var. Red.`)
cuped_vr_4 <- all_results %>% filter(Method == "CUPED", Window == "4-week") %>% pull(`Var. Red.`)
fd_vr_1 <- all_results %>% filter(Method == "First Diff.", Window == "1-week") %>% pull(`Var. Red.`)
fd_vr_4 <- all_results %>% filter(Method == "First Diff.", Window == "4-week") %>% pull(`Var. Red.`)
```

The pre-post correlation for existing users increases from `r round(rho_1, 3)` (1-week window) to `r round(rho_2, 3)` (2 weeks) and `r round(rho_4, 3)` (4 weeks), because averaging over more weeks smooths out noise and yields a more stable estimate of each user's baseline. The improvement from 1 to 2 weeks is larger than from 2 to 4 weeks, consistent with diminishing returns reported by Faire (Ritchie & Wong, 2021). Since CUPED's theoretical variance reduction equals $\rho^2$, these correlations imply that extending the covariate window from 1 to 4 weeks raises variance reduction from roughly `r round(rho_1^2 * 100)`% to `r round(rho_4^2 * 100)`%.

We compare four estimators: naive difference-in-means, first differences ($Y_{post} - X_{pre}$, equivalent to CUPED with $\theta$ forced to 1), CUPED (data-driven optimal $\hat{\theta}$), and OLS regression adjustment. For new users with missing pre-experiment data, we apply the Booking.com COALESCE approach and leave their outcome unadjusted (Jackson, 2018). Full numerical results are reported in Appendix B (Tables B1 and B2).

### Results

CUPED and OLS regression adjustment produce identical point estimates, standard errors, and confidence intervals across all window lengths. This follows from the fact that the two methods are mathematically equivalent when a single covariate is used: CUPED's $\hat{\theta} = \text{Cov}(Y, X)/\text{Var}(X)$ is the OLS slope on $X$, so the adjusted outcomes are the same.

First differences performs slightly worse than CUPED because it fixes $\theta = 1$, while the data-driven $\hat{\theta}$ ranges from about 0.69 (1-week) to 0.90 (4-week). The gap is largest at short windows: `r fd_vr_1` variance reduction for first differences versus `r cuped_vr_1` for CUPED at 1 week. As the window lengthens and $\hat{\theta}$ approaches 1, the two methods converge (`r fd_vr_4` versus `r cuped_vr_4` at 4 weeks).

Naive DM fails to detect the 0.10 hrs/week treatment effect at any window length because its confidence interval always includes zero. CUPED and first differences both reach significance at the 2-week and 4-week windows. When new users (20% of the sample) are included via the COALESCE approach, CUPED with a 4-week window still achieves `r cuped_all_vr` variance reduction at the population level, compared to `r cuped_vr_4` for existing users alone. New users in isolation have a standard error of `r new_se`, roughly twice that of the full sample (`r ex_naive_se`), confirming the Netflix finding that CUPED provides no benefit for users without historical data (Xie & Aurisset, 2016).

\FloatBarrier

```{r poc-ci-plot, echo=FALSE, fig.width=6, fig.height=4, fig.align="center"}
plot_data <- all_results %>%
  filter(Method %in% c("Naive DM", "First Diff.", "CUPED")) %>%
  mutate(
    CI_lo = as.numeric(str_extract(`95% CI`, "(?<=\\[)[\\d.-]+")),
    CI_hi = as.numeric(str_extract(`95% CI`, "(?<=, )[\\d.-]+(?=\\])"))
  ) %>%
  rename(Est = Estimate)

plot_data$Window <- factor(plot_data$Window, levels = c("1-week", "2-week", "4-week"))
plot_data$Method <- factor(plot_data$Method, levels = c("Naive DM", "First Diff.", "CUPED"))

ggplot(plot_data, aes(x = Window, y = Est, color = Method)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -0.005, ymax = 0.005,
           fill = "gray90", alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_hline(yintercept = tau, linetype = "dotted", color = "red", alpha = 0.7) +
  geom_pointrange(aes(ymin = CI_lo, ymax = CI_hi),
                  position = position_dodge(width = 0.4), size = 0.7, linewidth = 0.9) +
  annotate("text", x = 0.6, y = tau + 0.015, label = paste0("True effect = ", tau),
           color = "red", size = 2.8, hjust = 0) +
  annotate("text", x = 0.6, y = -0.015, label = "zero",
           color = "gray40", size = 2.5, hjust = 0) +
  scale_color_manual(values = c("Naive DM" = "gray50", "First Diff." = "orange3", "CUPED" = "steelblue")) +
  labs(y = "Estimated Treatment Effect (hrs/week)",
       x = "Pre-Experiment Window Length") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank())
```

*Figure 1: 95% confidence intervals by method and window length (existing users). OLS Reg. Adj. omitted (identical to CUPED with single covariate).*

---

# Appendix

## [A] Simulation Code

```{r appendix-code, eval=FALSE}
library(tidyverse)
set.seed(132)

# -- Parameters --
N <- 50000
new_user_frac <- 0.20
log_mean <- log(7)
log_sd <- 0.7
noise_sd <- 4.0
tau <- 0.10
n_pre_weeks <- 4

# -- Generate Users --
n_new <- round(N * new_user_frac)
n_existing <- N - n_new

users <- tibble(
  user_id = 1:N,
  is_new = c(rep(TRUE, n_new), rep(FALSE, n_existing)),
  baseline_mu = rlnorm(N, meanlog = log_mean, sdlog = log_sd),
  treatment = sample(c(0L, 1L), N, replace = TRUE)
)

# -- Generate Weekly Streaming Hours --
generate_week <- function(baseline_mu, treatment_indicator = NULL, tau_val = 0) {
  n <- length(baseline_mu)
  noise <- rnorm(n, mean = 0, sd = noise_sd)
  hours <- baseline_mu + noise
  if (!is.null(treatment_indicator)) {
    hours <- hours + tau_val * treatment_indicator
  }
  return(pmax(hours, 0))
}

pre_weeks <- map(1:n_pre_weeks, ~ generate_week(users$baseline_mu))
names(pre_weeks) <- paste0("pre_week_", 1:n_pre_weeks)
post_week <- generate_week(users$baseline_mu, users$treatment, tau)

# -- Assemble Dataset --
sim_data <- users %>%
  bind_cols(as_tibble(pre_weeks)) %>%
  mutate(
    post_week = post_week,
    pre_1wk = pre_week_4,
    pre_2wk = (pre_week_3 + pre_week_4) / 2,
    pre_4wk = (pre_week_1 + pre_week_2 + pre_week_3 + pre_week_4) / 4
  ) %>%
  mutate(
    pre_1wk = if_else(is_new, NA_real_, pre_1wk),
    pre_2wk = if_else(is_new, NA_real_, pre_2wk),
    pre_4wk = if_else(is_new, NA_real_, pre_4wk)
  )

# -- CUPED Adjustment (Booking.com COALESCE for missing data) --
cuped_adjust <- function(y, x) {
  has_x <- !is.na(x)
  theta <- cov(y[has_x], x[has_x]) / var(x[has_x])
  x_mean <- mean(x[has_x])
  y_adj <- y
  y_adj[has_x] <- y[has_x] - theta * (x[has_x] - x_mean)
  return(list(y_adj = y_adj, theta = theta))
}

# -- Four-Method Comparison --
run_test_all <- function(data, pre_col, label) {
  y <- data$post_week
  x <- data[[pre_col]]
  w <- data$treatment
  has_x <- !is.na(x)

  y_t <- y[w == 1]; y_c <- y[w == 0]
  n_t <- length(y_t); n_c <- length(y_c)

  # Naive DM
  naive_est <- mean(y_t) - mean(y_c)
  naive_se <- sqrt(var(y_t) / n_t + var(y_c) / n_c)

  # First Differences (theta = 1)
  fd <- y; fd[has_x] <- y[has_x] - x[has_x]
  fd_t <- fd[w == 1]; fd_c <- fd[w == 0]
  fd_est <- mean(fd_t) - mean(fd_c)
  fd_se <- sqrt(var(fd_t) / n_t + var(fd_c) / n_c)

  # CUPED (optimal theta)
  res <- cuped_adjust(y, x)
  ya <- res$y_adj
  ya_t <- ya[w == 1]; ya_c <- ya[w == 0]
  cuped_est <- mean(ya_t) - mean(ya_c)
  cuped_se <- sqrt(var(ya_t) / n_t + var(ya_c) / n_c)

  # OLS Regression Adjustment
  df_ols <- tibble(y = y, w = w, x = x) %>% filter(!is.na(x))
  mod <- lm(y ~ w + x, data = df_ols)
  ols_est <- coef(mod)["w"]
  ols_se <- summary(mod)$coefficients["w", "Std. Error"]

  cat(sprintf("[%s]\n", label))
  cat(sprintf("  Naive DM:     est=%.4f SE=%.4f VR=  0.0%%\n", naive_est, naive_se))
  cat(sprintf("  First Diff:   est=%.4f SE=%.4f VR=%5.1f%%\n",
              fd_est, fd_se, (1 - fd_se^2/naive_se^2)*100))
  cat(sprintf("  CUPED:        est=%.4f SE=%.4f VR=%5.1f%%\n",
              cuped_est, cuped_se, (1 - cuped_se^2/naive_se^2)*100))
  cat(sprintf("  OLS Reg Adj:  est=%.4f SE=%.4f VR=%5.1f%%\n\n",
              ols_est, ols_se, (1 - ols_se^2/naive_se^2)*100))
}

existing <- sim_data %>% filter(!is_new)
run_test_all(existing, "pre_1wk", "Existing, 1-week")
run_test_all(existing, "pre_2wk", "Existing, 2-week")
run_test_all(existing, "pre_4wk", "Existing, 4-week")
run_test_all(sim_data, "pre_1wk", "All users, 1-week")
run_test_all(sim_data, "pre_4wk", "All users, 4-week")
```

## [B] Detailed Simulation Results

```{r appendix-tables, echo=FALSE}
# Recompute (these chunks run independently of Section 7 chunks)
existing_app <- sim_data %>% filter(!is_new)

all_results_app <- bind_rows(
  run_test_all(existing_app, "pre_1wk", "1-week"),
  run_test_all(existing_app, "pre_2wk", "2-week"),
  run_test_all(existing_app, "pre_4wk", "4-week")
)

rho_1_app <- cor(existing_app$pre_1wk, existing_app$post_week)
rho_2_app <- cor(existing_app$pre_2wk, existing_app$post_week)
rho_4_app <- cor(existing_app$pre_4wk, existing_app$post_week)

# Table B0: Pre-post correlation
tibble(
  `Pre-Experiment Window` = c("1 week", "2 weeks", "4 weeks"),
  `Correlation (rho)` = round(c(rho_1_app, rho_2_app, rho_4_app), 3),
  `Theoretical VR` = paste0(round(c(rho_1_app, rho_2_app, rho_4_app)^2 * 100, 1), "%")
) %>% knitr::kable(caption = "Table B0: Pre-post correlation and theoretical variance reduction by window length (existing users).")

# Table B1: SE and Variance Reduction
all_results_app %>%
  select(Window, Method, SE, `Var. Red.`) %>%
  pivot_wider(names_from = Method, values_from = c(SE, `Var. Red.`)) %>%
  select(Window,
         `Naive SE` = `SE_Naive DM`,
         `FD SE` = `SE_First Diff.`,
         `CUPED SE` = `SE_CUPED`,
         `OLS SE` = `SE_OLS Reg. Adj.`,
         `FD VR` = `Var. Red._First Diff.`,
         `CUPED VR` = `Var. Red._CUPED`,
         `OLS VR` = `Var. Red._OLS Reg. Adj.`) %>%
  knitr::kable(caption = "Table B1: Standard errors and variance reduction by method (existing users, true tau = 0.10).")

# Table B2: Estimates and CIs
all_results_app %>%
  filter(Method != "OLS Reg. Adj.") %>%
  mutate(Result = sprintf("%.3f %s %s", Estimate, `95% CI`, ifelse(Sig. == "Yes", "*", ""))) %>%
  select(Window, Method, Result) %>%
  pivot_wider(names_from = Method, values_from = Result) %>%
  knitr::kable(caption = "Table B2: Estimates and 95% CIs (existing users). * = significant at 5%. OLS identical to CUPED, omitted.")
```
