---
title: "CUPED Simulation: Streaming Platform A/B Test"
author: "ECMA 31370 White Paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

## 1. Overview

We simulate an A/B test on a mid-tier streaming platform testing a new recommendation algorithm. The goal is to demonstrate how CUPED (Controlled-experiment Using Pre-Experiment Data) reduces the variance of the treatment effect estimator.

The simulation is designed to reproduce key findings from the literature:

- **Netflix (Xie & Aurisset, 2016):** ~40% variance reduction for existing users; negligible benefit for new users
- **Faire (Ritchie & Wong, 2021):** longer pre-experiment windows yield more variance reduction with diminishing returns
- **Booking.com (Jackson, 2018):** users with missing pre-experiment data can be left unadjusted (COALESCE approach)

## 2. Data Generating Process

Each user $i$ has a stable baseline streaming rate $\mu_i$ drawn from a lognormal distribution (right-skewed, matching the empirical shape of streaming hours). Weekly streaming hours are:

$$\text{hours}_{it} = \max\left(0,\; \mu_i + \tau \cdot W_i + \varepsilon_{it}\right)$$

where $W_i \in \{0, 1\}$ is the treatment indicator and $\varepsilon_{it} \sim N(0, \sigma_\varepsilon^2)$ is week-to-week noise. The $\max(0, \cdot)$ floor ensures non-negative hours. Light users with small $\mu_i$ naturally produce zero-hours weeks when noise pushes them below zero. Because the same users tend to be light across all weeks, these zeros are correlated in pre and post periods, preserving realistic pre-post correlation.

```{r parameters}
set.seed(42)

N <- 50000               # total users
new_user_frac <- 0.20    # 20% new users (no history)

log_mean <- log(7)       # baseline median ~7 hrs/week
log_sd <- 0.7            # right-skewed spread

noise_sd <- 5.0          # week-to-week noise (tuned for rho ~ 0.70)

tau <- 0.10              # treatment effect: +0.10 hrs/week (~1% relative lift)

n_pre_weeks <- 4         # generate 4 weeks of pre-experiment data
```

**Parameter choices:**

| Parameter | Value | Rationale |
|-----------|-------|-----------|
| N | 50,000 | Mid-tier platform experiment scale |
| New user fraction | 20% | Growing platform |
| Baseline $\mu_i$ | LogNormal(log 7, 0.7) | Median 7 hrs/week, right-skewed to 30+ |
| Noise $\sigma_\varepsilon$ | 5.0 | Tuned so $\rho \approx 0.70$ for 1-week window |
| Treatment $\tau$ | 0.10 hrs/week | Small enough that naive method struggles |

## 3. Generate Users

```{r generate-users}
n_new <- round(N * new_user_frac)
n_existing <- N - n_new

users <- tibble(
  user_id = 1:N,
  is_new = c(rep(TRUE, n_new), rep(FALSE, n_existing)),
  baseline_mu = rlnorm(N, meanlog = log_mean, sdlog = log_sd),
  treatment = sample(c(0L, 1L), N, replace = TRUE)
)

cat(sprintf("Treatment: %d | Control: %d\n",
            sum(users$treatment == 1), sum(users$treatment == 0)))
cat(sprintf("New: %d | Existing: %d\n", n_new, n_existing))
```

## 4. Generate Weekly Streaming Hours

```{r generate-weeks}
generate_week <- function(baseline_mu, treatment_indicator = NULL, tau_val = 0) {
  n <- length(baseline_mu)
  noise <- rnorm(n, mean = 0, sd = noise_sd)
  hours <- baseline_mu + noise
  if (!is.null(treatment_indicator)) {
    hours <- hours + tau_val * treatment_indicator
  }
  return(pmax(hours, 0))
}

# 4 weeks pre-experiment (no treatment effect)
pre_weeks <- map(1:n_pre_weeks, ~ generate_week(users$baseline_mu))
names(pre_weeks) <- paste0("pre_week_", 1:n_pre_weeks)

# 1 week experiment period (treatment group gets tau)
post_week <- generate_week(users$baseline_mu, users$treatment, tau)
```

## 5. Assemble Dataset

We construct three pre-experiment covariates at different window lengths: 1-week (most recent week only), 2-week average, and 4-week average. For new users, all pre-experiment columns are set to `NA`.

```{r assemble}
sim_data <- users %>%
  bind_cols(as_tibble(pre_weeks)) %>%
  mutate(
    post_week = post_week,
    pre_1wk = pre_week_4,
    pre_2wk = (pre_week_3 + pre_week_4) / 2,
    pre_4wk = (pre_week_1 + pre_week_2 + pre_week_3 + pre_week_4) / 4
  ) %>%
  mutate(
    pre_1wk = if_else(is_new, NA_real_, pre_1wk),
    pre_2wk = if_else(is_new, NA_real_, pre_2wk),
    pre_4wk = if_else(is_new, NA_real_, pre_4wk)
  )

head(sim_data %>% select(user_id, is_new, treatment, pre_1wk, pre_4wk, post_week), 10)
```

## 6. Explore the Data

### 6.1 Distribution of Streaming Hours

```{r dist-plot, fig.width=8, fig.height=4}
ggplot(sim_data, aes(x = post_week)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white", alpha = 0.8) +
  labs(title = "Distribution of Weekly Streaming Hours (Experiment Period)",
       x = "Streaming Hours / Week", y = "Count") +
  theme_minimal()
```

```{r dist-summary}
cat(sprintf("Mean:   %.2f hrs/week\n", mean(sim_data$post_week)))
cat(sprintf("Median: %.2f hrs/week\n", median(sim_data$post_week)))
cat(sprintf("SD:     %.2f\n", sd(sim_data$post_week)))
cat(sprintf("Zero-hours users: %.1f%%\n", mean(sim_data$post_week == 0) * 100))
```

### 6.2 Pre-Post Correlation

This is the key number that determines CUPED effectiveness. Higher $\rho$ means more variance reduction ($\text{VR} = \rho^2$).

```{r pre-post-corr}
existing <- sim_data %>% filter(!is_new)

rho_1 <- cor(existing$pre_1wk, existing$post_week)
rho_2 <- cor(existing$pre_2wk, existing$post_week)
rho_4 <- cor(existing$pre_4wk, existing$post_week)

tibble(
  Window = c("1 week", "2 weeks", "4 weeks"),
  rho = c(rho_1, rho_2, rho_4),
  `Theoretical VR` = paste0(round(c(rho_1, rho_2, rho_4)^2 * 100, 1), "%")
) %>% knitr::kable()
```

Longer windows increase $\rho$ because averaging over multiple weeks smooths out the noise, producing a more stable estimate of each user's baseline. This matches Faire's finding that variance reduction increases with window length but with diminishing returns.

### 6.3 Pre vs Post Scatter Plot

```{r scatter, fig.width=8, fig.height=4}
existing %>%
  sample_n(5000) %>%
  ggplot(aes(x = pre_4wk, y = post_week)) +
  geom_point(alpha = 0.15, size = 0.8, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 0.8) +
  labs(title = sprintf("Pre vs Post Streaming Hours (4-week window, rho = %.3f)", rho_4),
       x = "Pre-Experiment (4-week avg, hrs/week)",
       y = "Experiment Period (hrs/week)") +
  theme_minimal()
```

The strong linear relationship is what CUPED exploits: by regressing out the pre-experiment baseline, we remove predictable user-level variation and isolate the treatment effect.

## 7. CUPED Implementation

The CUPED-adjusted outcome is:

$$Y_i^{cv} = Y_i - \hat{\theta}(X_i - \bar{X})$$

where $\hat{\theta} = \text{Cov}(Y, X) / \text{Var}(X)$ and $X_i$ is the pre-experiment covariate.

For users with missing $X_i$ (new users), we apply the Booking.com COALESCE approach: leave their metric unadjusted. This is equivalent to imputing missing covariates with the sample mean, which zeroes out the adjustment term.

```{r cuped-function}
cuped_adjust <- function(y, x) {
  has_x <- !is.na(x)
  theta <- cov(y[has_x], x[has_x]) / var(x[has_x])
  x_mean <- mean(x[has_x])
  y_adj <- y
  y_adj[has_x] <- y[has_x] - theta * (x[has_x] - x_mean)
  return(list(y_adj = y_adj, theta = theta))
}
```

## 8. Results: Naive vs CUPED

```{r test-function}
run_test <- function(data, pre_col, label) {
  y <- data$post_week
  x <- data[[pre_col]]
  w <- data$treatment

  # naive difference-in-means
  y_t <- y[w == 1]; y_c <- y[w == 0]
  n_t <- length(y_t); n_c <- length(y_c)
  naive_est <- mean(y_t) - mean(y_c)
  naive_se <- sqrt(var(y_t) / n_t + var(y_c) / n_c)
  naive_ci <- naive_est + c(-1, 1) * 1.96 * naive_se
  naive_sig <- (naive_ci[1] > 0) | (naive_ci[2] < 0)

  # CUPED
  res <- cuped_adjust(y, x)
  ya <- res$y_adj
  ya_t <- ya[w == 1]; ya_c <- ya[w == 0]
  cuped_est <- mean(ya_t) - mean(ya_c)
  cuped_se <- sqrt(var(ya_t) / n_t + var(ya_c) / n_c)
  cuped_ci <- cuped_est + c(-1, 1) * 1.96 * cuped_se
  cuped_sig <- (cuped_ci[1] > 0) | (cuped_ci[2] < 0)

  vr <- 1 - cuped_se^2 / naive_se^2

  tibble(
    Label = label,
    `Naive Est` = naive_est,
    `Naive SE` = naive_se,
    `Naive CI` = sprintf("[%.4f, %.4f]", naive_ci[1], naive_ci[2]),
    `Naive Sig` = ifelse(naive_sig, "Yes", "No"),
    `CUPED Est` = cuped_est,
    `CUPED SE` = cuped_se,
    `CUPED CI` = sprintf("[%.4f, %.4f]", cuped_ci[1], cuped_ci[2]),
    `CUPED Sig` = ifelse(cuped_sig, "Yes", "No"),
    `Var Reduction` = sprintf("%.1f%%", vr * 100)
  )
}
```

### 8.1 Existing Users Only

This is the most informative comparison. New users are excluded so that CUPED operates on the full subsample.

```{r results-existing}
bind_rows(
  run_test(existing, "pre_1wk", "1-week window"),
  run_test(existing, "pre_2wk", "2-week window"),
  run_test(existing, "pre_4wk", "4-week window")
) %>% knitr::kable()
```

**Key finding:** The naive difference-in-means fails to detect the treatment effect (CI includes zero). CUPED with a 2-week or 4-week window successfully detects it. This demonstrates that CUPED does not change the point estimate (unbiased), but narrows the confidence interval enough to reach significance.

The 1-week window is not sufficient for CUPED to reach significance either, confirming the importance of window length.

### 8.2 All Users (Including New Users)

When new users are included, the Booking.com COALESCE approach is applied: new users are left unadjusted, existing users get the CUPED adjustment.

```{r results-all}
bind_rows(
  run_test(sim_data, "pre_1wk", "1-week window"),
  run_test(sim_data, "pre_2wk", "2-week window"),
  run_test(sim_data, "pre_4wk", "4-week window")
) %>% knitr::kable()
```

### 8.3 New Users Only

```{r results-new}
new_users <- sim_data %>% filter(is_new)
y_t <- new_users$post_week[new_users$treatment == 1]
y_c <- new_users$post_week[new_users$treatment == 0]
est <- mean(y_t) - mean(y_c)
se <- sqrt(var(y_t) / length(y_t) + var(y_c) / length(y_c))
ci <- est + c(-1, 1) * 1.96 * se

tibble(
  `Estimate` = est,
  `SE` = se,
  `95% CI` = sprintf("[%.4f, %.4f]", ci[1], ci[2]),
  `Significant` = ifelse((ci[1] > 0) | (ci[2] < 0), "Yes", "No")
) %>% knitr::kable()
```

New users have no pre-experiment data, so CUPED cannot be applied. The naive estimator alone has very high variance (SE is roughly 2x that of existing users), making it impossible to detect the small treatment effect. This confirms the Netflix finding that CUPED provides no benefit for users without historical data.

## 9. Visualize the Variance Reduction

```{r ci-plot, fig.width=8, fig.height=5}
# build comparison data for existing users
results <- bind_rows(
  run_test(existing, "pre_1wk", "1-week"),
  run_test(existing, "pre_2wk", "2-week"),
  run_test(existing, "pre_4wk", "4-week")
)

plot_data <- bind_rows(
  tibble(
    Window = results$Label,
    Method = "Naive",
    Est = results$`Naive Est`,
    CI_lo = as.numeric(str_extract(results$`Naive CI`, "(?<=\\[)[\\d.-]+")),
    CI_hi = as.numeric(str_extract(results$`Naive CI`, "(?<=, )[\\d.-]+(?=\\])"))
  ),
  tibble(
    Window = results$Label,
    Method = "CUPED",
    Est = results$`CUPED Est`,
    CI_lo = as.numeric(str_extract(results$`CUPED CI`, "(?<=\\[)[\\d.-]+")),
    CI_hi = as.numeric(str_extract(results$`CUPED CI`, "(?<=, )[\\d.-]+(?=\\])"))
  )
)

plot_data$Window <- factor(plot_data$Window, levels = c("1-week", "2-week", "4-week"))
plot_data$Method <- factor(plot_data$Method, levels = c("Naive", "CUPED"))

ggplot(plot_data, aes(x = Window, y = Est, color = Method)) +
  geom_pointrange(aes(ymin = CI_lo, ymax = CI_hi),
                  position = position_dodge(width = 0.3), size = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_hline(yintercept = tau, linetype = "dotted", color = "red", alpha = 0.6) +
  annotate("text", x = 3.4, y = tau, label = sprintf("True tau = %.2f", tau),
           color = "red", size = 3, alpha = 0.7) +
  scale_color_manual(values = c("Naive" = "gray50", "CUPED" = "steelblue")) +
  labs(title = "Naive vs CUPED: 95% Confidence Intervals (Existing Users)",
       y = "Estimated Treatment Effect (hrs/week)",
       x = "Pre-Experiment Window Length") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

The plot shows that CUPED consistently produces narrower confidence intervals than the naive method. With a 2-week or 4-week window, CUPED's confidence interval clears zero (statistically significant), while the naive interval does not.

## 10. Save Data

```{r save}
output <- sim_data %>%
  select(user_id, is_new, treatment, pre_1wk, pre_2wk, pre_4wk, post_week)

write_csv(output, "/tmp/caus_whitepaper/code/simulated_streaming_data.csv")
cat(sprintf("Saved: %d rows x %d columns\n", nrow(output), ncol(output)))
```
